The evolution of gene expression in the term placenta of viviparous mammals

### 2016-03-22 manuscript/manuscript.Rnw
add the GSE numbers to the supplement and the manuscript
add PE specific genes

All of the code used to produce every analysis presented in this paper
is available at
\url{https://github.com/uiuc-cgm/placenta-viviparous.git}. All of the
sequences used are deposited in SRA, and can be obtained using
\DLA{add GSE label for the experiment}.

@@

igf.exp.long$type <- "IGF Related"
gal.exp.long$type <- "Galectin"
hor.exp.long$type <- "Hormone"

### scalerange <- range(c(igf.exp.long$fpkm,gal.exp.long$fpkm,hor.exp.long$fpkm),na.rm=TRUE)
### gradientends <- scalerange + rep(c(0,100,200), each=2)
### colorends <- c("white", "red", "white", "green", "white", "blue")
### 
### gal.exp.long$fpkm <- gal.exp.long$fpkm+100
### hor.exp.long$fpkm <- hor.exp.long$fpkm+200

combined.long <-
    rbind(igf.exp.long,
          gal.exp.long,
          hor.exp.long)
combined.long$species <-
    factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(combined.long$species)),
           levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(names(species.names))))

common.plot.options <- 
    list(geom_tile(aes(fill = fpkm), colour = "white"),
##         scale_fill_gradientn(colours = colorends, values = rescale(gradientends)),
         scale_fill_gradient(low = grad.def[["low"]],
                             high = grad.def[["high"]],
                             space = grad.def[["space"]],
                             na.value = grad.def[["na.value"]],
                             guide=guide_colorbar(title=expression(log[10](FPKM)))),
         scale_x_discrete("", expand = c(0, 0)),
         scale_y_discrete("", expand = c(0, 0)),
         theme_grey(base_size = 9),
         theme(legend.position = "top",
               axis.ticks = element_blank(),
               panel.background = element_blank(),
               axis.text.y = element_text(face="italic"),
               axis.text.x = element_text(angle = 300, hjust = 0, vjust=1,face="italic")),
         facet_grid(type~.,scale="free_y",space="free_y")
         )
print(ggplot(combined.long, aes(y=gene, x=species))
      + common.plot.options
      )
@ 
\caption{Expression of Hormone Genes, Galectins, and IGF related genes}
\label{fig:hormone_genes}
\end{figure}

>>>>

All of the code used to produce every analysis presented in this paper
is available at
\url{https://github.com/uiuc-cgm/placenta-viviparous.git}. All of the
sequences used are deposited in SRA, and can be obtained under the
series
\href{http://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE79121}{GSE79121}.

@@

### Pre-eclampsia genes FSTL3 EPAS1 ADAM12 GPC3 TGFB1 PAPPA

pe.expression.nozero <-
    gene.expression.nozero(c("ADAM12","CRH","KISS1","PAPPA"))

pe.exp.long <- melt(log10(pe.expression.nozero+1)[rev(1:nrow(pe.expression.nozero)),])
colnames(pe.exp.long)[1:3] <- c("gene","species","fpkm")

igf.exp.long$type <- "IGF Related"
gal.exp.long$type <- "Galectin"
hor.exp.long$type <- "Hormone"
pe.exp.long$type <- "PE"

combined.long <-
    rbind(igf.exp.long,
          gal.exp.long,
          hor.exp.long,
          pe.exp.long
          )
combined.long$species <-
    factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(combined.long$species)),
           levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(names(species.names))))

common.plot.options <- 
    list(geom_tile(aes(fill = fpkm), colour = "white"),
         scale_fill_gradient(low = grad.def[["low"]],
                             high = grad.def[["high"]],
                             space = grad.def[["space"]],
                             na.value = grad.def[["na.value"]],
                             guide=guide_colorbar(title=expression(log[10](FPKM)))),
         scale_x_discrete("", expand = c(0, 0)),
         scale_y_discrete("", expand = c(0, 0)),
         theme_grey(base_size = 9),
         theme(legend.position = "top",
               axis.ticks = element_blank(),
               panel.background = element_blank(),
               axis.text.y = element_text(face="italic"),
               axis.text.x = element_text(angle = 300, hjust = 0, vjust=1,face="italic")),
         facet_grid(type~.,scale="free_y",space="free_y")
         )
print(ggplot(combined.long, aes(y=gene, x=species))
      + common.plot.options
      )
@ 
\caption{Expression of 1:1 Human orthologs of Hormone Genes, Galectins,
  genes implicated in pre-eclampsia (PE), and \textit{IGF}-related genes}
\label{fig:hormone_genes}
\end{figure}
